---
title: "Getting Started"
author: "kevinoliva26"
date: "10/16/2020"
output: html_document
---

Normalmente se comienza cargando las librerias
Ejemplo de una grafica
```{r}
library(tidyverse)
library(dslabs)
data(murders)

murders %>%
    ggplot(aes(population, total, label=abb, color=region)) +
    geom_label()
```

To define a variable, we may use the assignment symbol, <-
```{r}
# assigning values to variables
a <-2
b <--1
c <--4

# solving the quadratic equation
(-b + sqrt(b^2 - 4*a*c))/(2*a)
(-b - sqrt(b^2 - 4*a*c))/(2*a)
```

The ls() function shows the names of the objects saved in your workspace.
```{r}
ls()
```

To access help files, we may use the help function, help(function_name), or write the question mark followed by the function name, ?function_name.

The args() function also shows the arguments a function needs.

To specify arguments, we use the equals sign. 

The function class() helps us determine the type of an object.

To access data from columns of a data frame, we use the dollar sign symbol, $, which is called the accessor.

Factors are useful for storing categorical data, and are more memory efficient than storing characters.
```{r}
# loading the dslabs package and the murders dataset
library(dslabs)
data(murders)

# determining that the murders dataset is of the "data frame" class
class(murders)
# finding out more about the structure of the object
str(murders)
# showing the first 6 lines of the dataset
head(murders)

# using the accessor operator to obtain the population column
murders$population
# displaying the variable names in the murders dataset
names(murders)
# determining how many entries are in a vector
pop <- murders$population
length(pop)
# vectors can be of class numeric and character
class(pop)
class(murders$state)

# logical vectors are either TRUE or FALSE
z <- 3 == 2
z
class(z)

# factors are another type of class
class(murders$region)
# obtaining the levels of a factor
levels(murders$region)
```


The function table takes a vector as input and returns the frequency of each unique element in the vector.
```{r}
# Write one line of code to show the number of states per region
table(murders$region)
```

