---
title: 'Section 2: Vectors, Sorting'
author: "kevinoliva26"
date: "10/16/2020"
output: html_document
---
## Vectors

The function c(), which stands for concatenate, is useful for creating vectors.
Another useful function for creating vectors is the seq() function, which generates sequences.
Subsetting lets us access specific parts of a vector by using square brackets to access elements of a vector.

```{r}
# We may create vectors of class numeric or character with the concatenate function
codes <- c(380, 124, 818)
country <- c("italy", "canada", "egypt")

# We can also name the elements of a numeric vector
# Note that the two lines of code below have the same result
codes <- c(italy = 380, canada = 124, egypt = 818)
codes <- c("italy" = 380, "canada" = 124, "egypt" = 818)

# We can also name the elements of a numeric vector using the names() function
codes <- c(380, 124, 818)
country <- c("italy","canada","egypt")
names(codes) <- country

# Using square brackets is useful for subsetting to access specific elements of a vector
codes[2]
codes[c(1,3)]
codes[1:2]

# If the entries of a vector are named, they may be accessed by referring to their name
codes["canada"]
codes[c("egypt","italy")]
```

In general, coercion is an attempt by R to be flexible with data types by guessing what was meant when an entry does not match the expected. For example, when defining x as

```{r}
x <- c(1, "canada", 3)
x
```

The function *as.character()* turns numbers into characters.
The function *as.numeric()* turns characters into numbers.
In R, missing data is assigned the value NA.

## Sorting

The function *sort()* sorts a vector in increasing order.
The function *order()* produces the indices needed to obtain the sorted vector, e.g. a result of  2 3 1 5 4 means the sorted vector will be produced by listing the 2nd, 3rd, 1st, 5th, and then 4th item of the original vector.
The function *rank()* gives us the ranks of the items in the original vector.
The function *max()* returns the largest value, while *which.max()* returns the index of the largest value. The functions *min()* and *which.min()* work similarly for minimum values.

```{r}
library(dslabs)
data(murders)
sort(murders$total)

x <- c(31, 4, 15, 92, 65)
x
sort(x)    # puts elements in order

index <- order(x)    # returns index that will put x in order
x[index]    # rearranging by this index puts elements in order
order(x)

murders$state[1:10]
murders$abb[1:10]

index <- order(murders$total)
murders$abb[index]    # order abbreviations by total murders

max(murders$total)    # highest number of total murders
i_max <- which.max(murders$total)    # index with highest number of murders
murders$state[i_max]    # state name with highest number of total murders

x <- c(31, 4, 15, 92, 65)
x
rank(x)    # returns ranks (smallest to largest)
```

Funcion *is.na()*
```{r}
# Using new dataset 
library(dslabs)
data(na_example)

# Checking the structure 
str(na_example)

# Find out the mean of the entire dataset 
mean(na_example)

# Use is.na to create a logical index ind that tells which entries are NA
ind <- is.na(na_example)

# Determine how many NA ind has using the sum function
sum(ind)
```


##Vector Arithmetic

```{r}
# The name of the state with the maximum population is found by doing the following
murders$state[which.max(murders$population)]

# how to obtain the murder rate
murder_rate <- murders$total / murders$population * 100000

# ordering the states by murder rate, in decreasing order
murders$state[order(murder_rate, decreasing=TRUE)]
```


## Section 2 Assessment

```{r}
x <- c(2, 43, 27, 96, 18)
sort(x)
order(x)
rank(x)

name <- c("Mandi", "Amy", "Nicole", "Olivia")
distance <- c(0.8, 3.1, 2.8, 4.0)
time <- c(10, 30, 40, 50)

time <- time/60
time

speed <- distance / time
speed
```









