---
title: 'Section 3: Indexing, Data Wrangling, Plots'
author: "kevinoliva26"
date: "10/17/2020"
output: html_document
---

## Indexing

Using the function sum()on a logical vector returns the number of entries that are true.
The logical operator *“&”* makes two logicals true only when they are both true.

```{r}
# defining murder rate as before
murder_rate <- murders$total / murders$population * 100000
# creating a logical vector that specifies if the murder rate in that state is less than or equal to 0.71
index <- murder_rate <= 0.71
# determining which states have murder rates less than or equal to 0.71
murders$state[index]
# calculating how many states have a murder rate less than or equal to 0.71
sum(index)

# creating the two logical vectors representing our conditions
west <- murders$region == "West"
safe <- murder_rate <= 1
# defining an index and identifying states with both conditions true
index <- safe & west
murders$state[index]
```

The function *which()* gives us the entries of a logical vector that are true.
The function *match()* looks for entries in a vector and returns the index needed to access them.
We use the function *%in%* if we want to know whether or not each element of a first vector is in a second vector.


```{r}
x <- c(FALSE, TRUE, FALSE, TRUE, TRUE, FALSE)
which(x)    # returns indices that are TRUE

# to determine the murder rate in Massachusetts we may do the following
index <- which(murders$state == "Massachusetts")
index
murder_rate[index]

# to obtain the indices and subsequent murder rates of New York, Florida, Texas, we do:
index <- match(c("New York", "Florida", "Texas"), murders$state)
index
murders$state[index]
murder_rate[index]

x <- c("a", "b", "c", "d", "e")
y <- c("a", "d", "f")
y %in% x

# to see if Boston, Dakota, and Washington are states
c("Boston", "Dakota", "Washington") %in% murders$state
```

## Basic Data Wrangling

The *dplyr* package is introduced in the context of the tidyverse, a collection of R packages
 
To change a data table by adding a new column, or changing an existing one, we use the *mutate()* function.
To filter the data by subsetting rows, we use the function *filter()*.
To subset the data by selecting specific columns, we use the *select()* function.
We can perform a series of operations by sending the results of one function to another function using the pipe operator, *%>%*.

```{r}

# installing and loading the dplyr package
library(dplyr)

# adding a column with mutate
library(dslabs)
data("murders")
murders <- mutate(murders, rate = total / population * 100000)

# subsetting with filter
filter(murders, rate <= 0.71)

# selecting columns with select
new_table <- select(murders, state, region, rate)

# using the pipe
murders %>% select(state, region, rate) %>% filter(rate <= 0.71)
```

We can use the *data.frame()* function to create data frames.
Formerly, the *data.frame()* function turned characters into factors by default. To avoid this, we could utilize the stringsAsFactors argument and set it equal to false. As of R 4.0, it is no longer necessary to include the *stringsAsFactors* argument, because R no longer turns characters into factors by default.
*nrow()* sirve para contar el numero de filas

```{r}
# creating a data frame with stringAsFactors = FALSE
grades <- data.frame(names = c("John", "Juan", "Jean", "Yao"), 
                     exam_1 = c(95, 80, 90, 85), 
                     exam_2 = c(90, 85, 85, 90),
                     stringsAsFactors = FALSE)
```

Ejemplo de %>%

```{r}
# Loading the libraries
library(dplyr)
data(murders)

# Create new data frame called my_states (with specifications in the instructions)
my_states <- murders %>% mutate(rate = total / population * 100000, rank = rank(-rate) ) %>% filter(region %in% c("Northeast","West") & rate <1) %>% select(state, rate, rank)
my_states
```

















