---
title: "Section 4 Programming Basics"
author: "kevinoliva26"
date: "11/3/2020"
output: html_document
---

##Basic Conditionals

The most common conditional expression in programming is an if-else statement, which has the form *"if [condition], perform [expression], else perform [alternative expression]"*.
The *ifelse()* function works similarly to an if-else statement, but it is particularly useful since it works on vectors by examining each element of the vector and returning a corresponding answer accordingly.
The *any()* function takes a vector of logicals and returns true if any of the entries are true.
The *all()* function takes a vector of logicals and returns true if all of the entries are true.

```{r}
# an example showing the general structure of an if-else statement
a <- 0
if(a!=0){
  print(1/a)
} else{
  print("No reciprocal for 0.")
}

# an example that tells us which states, if any, have a murder rate less than 0.5
library(dslabs)
data(murders)
murder_rate <- murders$total / murders$population*100000
ind <- which.min(murder_rate)
if(murder_rate[ind] < 0.5){
  print(murders$state[ind]) 
} else{
  print("No state has murder rate that low")
}

# changing the condition to < 0.25 changes the result
if(murder_rate[ind] < 0.25){
  print(murders$state[ind]) 
} else{
  print("No state has a murder rate that low.")
}

# the ifelse() function works similarly to an if-else conditional
a <- 0
ifelse(a > 0, 1/a, NA)

# the ifelse() function is particularly useful on vectors
a <- c(0,1,2,-4,5)
result <- ifelse(a > 0, 1/a, NA)

# the ifelse() function is also helpful for replacing missing values
data(na_example)
no_nas <- ifelse(is.na(na_example), 0, na_example) 
sum(is.na(no_nas))

# the any() and all() functions evaluate logical vectors
z <- c(TRUE, TRUE, FALSE)
any(z)
all(z)
```

##Functions

The R function called *function()* tells R you are about to define a new function.
Functions are objects, so must be assigned a variable name with the arrow operator.
The general way to define functions is: (1) decide the function name, which will be an object, (2) type function() with your function's arguments in parentheses, (3) write all the operations inside brackets.
Variables defined inside a function are not saved in the workspace.

```{r}
# example of defining a function to compute the average of a vector x
avg <- function(x){
  s <- sum(x)
  n <- length(x)
  s/n
}

# we see that the above function and the pre-built R mean() function are identical
x <- 1:100
identical(mean(x), avg(x))

# variables inside a function are not defined in the workspace
s <- 3
avg(1:10)
s

# the general form of a function
my_function <- function(VARIABLE_NAME){
  perform operations on VARIABLE_NAME and calculate VALUE
  VALUE
}

# functions can have multiple arguments as well as default values
avg <- function(x, arithmetic = TRUE){
  n <- length(x)
  ifelse(arithmetic, sum(x)/n, prod(x)^(1/n))
}
```


##For Loops

For-loops perform the same task over and over while changing the variable.  They let us define the range that our variable takes, and then changes the value with each loop and evaluates the expression every time inside the loop.
The general form of a for-loop is: *"For i in [some range], do operations"*.  This *i*changes across the range of values and the operations assume i is a value you're interested in computing on.
At the end of the loop, the value of i is the last value of the range.

```{r}
# creating a function that computes the sum of integers 1 through n
compute_s_n <- function(n){
  x <- 1:n
  sum(x)
}

# a very simple for-loop
for(i in 1:5){
  print(i)
}

# a for-loop for our summation
m <- 25
s_n <- vector(length = m) # create an empty vector
for(n in 1:m){
  s_n[n] <- compute_s_n(n)
}
s_n

# creating a plot for our summation function
n <- 1:m
plot(n, s_n)

# a table of values comparing our function to the summation formula
head(data.frame(s_n = s_n, formula = n*(n+1)/2))

# overlaying our function with the summation formula
plot(n, s_n)
lines(n, n*(n+1)/2)
```


##Other Functions

Functions that are typically used instead of for loops in R
are apply, sapply, tapply, and mapply.
These are part of what we call the apply family.
Other functions that are widely used are split, cut, quantile, reduce,
identical, unique, and many others.

##Section 4 Quiz

Write an ifelse() statement that returns 1 if the sex is Female and 2 if the sex is Male.

```{r}
library(dslabs)
data(heights)
m_or_f <- ifelse(heights$sex=="Male",2,1)
sum(m_or_f)
```

Write an ifelse() statement that takes the height column and returns the height if it is greater than 72 inches and returns 0 otherwise.

What is the mean of the resulting vector?

```{r}
library(dslabs)
data(heights)
mean(ifelse(heights$height>72,heights$height,0))
```

Write a function inches_to_ft that takes a number of inches x and returns the number of feet. One foot equals 12 inches.

What is inches_to_ft(144)?

How many individuals in the heights dataset have a height less than 5 feet?

```{r}
inches_to_ft <- function(x){
  x/12
}
inches_to_ft(144)

sum(inches_to_ft(heights)$height<5)
```

Given an integer x, the factorial of x is called x! and is the product of all integers up to and including x. The factorial() function computes factorials in R. For example, factorial(4) returns 4! = 4 × 3 × 2 × 1 = 24.

```{r}
# define a vector of length m
m <- 10
f_n <- vector(length = m)

# make a vector of factorials
for(n in 1:m){
f_n[n] <- factorial(n)
}

# inspect f_n
f_n
```

